Towards a Machine Learning approach for Aggregate Computing

### 2021-11-28 paper-2021-woa-ac-ml.tex
On execution model

\subsection{Computational Model}
An aggregate program execution consists of a continuously node-point execution of a \round{}, 
 which is the atomical part of AC. Each node can interact only within its neighbourhoods sharing the \export{}, namely the result produced by the evaluation of a \round{}.

The main \round{} steps consist of:
\begin{enumerate}
  \item execution context creation: nodes create the execution context collecting information from the sensors, 
  the neighbourhood messages, and the old program output;
  \item program evaluation: the system-wide aggregate program is evaluated against the local context created by the nodes. 
  This produces the \export{};
  \item export sharing: the export created will be shared within the current neighbourhoods perceived by the nodes;
  \item actuations: giving the export produced, nodes may perform some actuation (e.g. motor activations, steering, ...).
\end{enumerate}

Rounds execution is completely asynchronous, so do not exist a global clock to coordinate the aggregate. 
 This makes it possible to scale easily with the node number size. 
 This collective, proactive and periodical execution will eventually lead to the satisfaction of the system-wide specification.

>>>>

\subsection{Computational Model}
%
For an aggreate program to yield collective adaptive behaviour, 
 continuous computation and communication is needed.
%
An individual atomic step 
 involving context acquisition,
 computation of the program,
 and neighbourhood-based communication 
 is called a \round{}.
%
In detail, a \round{} run at some device consists of:
\begin{enumerate}
  \item \emph{context acquisition}: the device collects information from the sensors, 
  the messages from neighbours
  (including the device itself---to model state);
  \item \emph{program evaluation}: the aggregate program is evaluated against the acquired context,
  yielding an \export{}, namely the message to be sent to neighbours;
  \item \emph{export sharing}: the \export{} is sent to the neighbours;
  \item \emph{actuations}: the \export{} also includes information that can be used to drive actuators in the local node.
\end{enumerate}

Rounds execution is completely asynchronous, so there is no global clock or barrier to coordinate the aggregate. 
%
 This makes it possible to scale easily with the number of nodes. 
%
The combination of the logic of the aggregate program
 and 
 such a collective, proactive, and periodical execution
 can promote the pursuing of collective adaptive behaviour.
